\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Local Adaptive Optimization of Time Step}

### 2013-10-22 report/article.tex
Introduction and Previous Work shipped for review. Mass spring system initiated

\usepackage{amsmath,amssymb,amsfonts}
\providecommand{\abs}[1]{\left \lvert #1 \right \rvert}

@@

\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Local Adaptive
Optimization of Time step}} % Article title

@@

\section{Introduction}
\lettrine[nindent=0em,lines=3]{S} imulations of large hyper elastic materials
are very time consuming. When working with materials of varying density
of vertices/scale of mesh, the highest densities often make the most
strict bounds on size of the time step of the simulation in order to keep
the materials (and grids) stable and the discretization correct. in a
``standard'' simulation with a global time step size we need to compute
possibly unnecessarily many computations on the coarser parts of the material
caused by this bound. When furthermore having materials with a large amount
of somewhat equally distributed vertices and smaller patches of detailed
areas with higher density of vertices, it would be an advantage to be able
to perform large time steps for the majority of the material and smaller
steps for the local patches where vertices could cause instability. we call
the concept of varying the timestep local adaptive optimization of time step
(LAOT), in which local either refer to spatial locality or time locality. We
will in this report study LAOT in the simple 1 dimensional simulation of a
mass spring system in order to gain knowledge into the field for further use
in simulation of hyper elastic materials. We will first describe the simulation
problem and then go into the study of the application of LOAT herein.

@@

We will in this report consider the 1 dimensional problem of simulating a mass
spring system, which is usually (in 3 dimensions) used for simulating
cloth and hair etc. A mass spring system is a system consisting of a set of
vertices $V$ each having a specified mass $m$, force $F$ and velocity $u$ and
a set of springs $S$ with two endpoints $p_f \in V$ and $p_t \in V \setminus
p_f$. Each spring $s$ has a damping factor $d_s$, a stiffness constant $K_s$ and a rest
length $l_{s0}$ describing the spring's physical properties.

@@

\subsection{Spatial LAOT}
Atomic update

>>>>

\usepackage{todonotes}
\usepackage{amsmath,amssymb,amsfonts}
\providecommand{\abs}[1]{\left \lvert #1 \right \rvert}

@@

\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Local Adaptive
Optimization of Time Step}} % Article title

@@

\section{Introduction}
\lettrine[nindent=0em,lines=3]{S} imulations of large hyper elastic materials
can be very time consuming. When working with materials of varying density
of vertices/scale of mesh, the highest densities often make the most
strict bounds on size of the time step of the simulation in order to keep
the materials (and grids) stable and the discretization correct. in a
``standard'' simulation with a global time step size we need to compute
possibly unnecessarily many computations on the coarser parts of the material
caused by this bound. When furthermore having materials with a large amount
of somewhat equally distributed vertices and smaller patches of detailed
areas with higher density of vertices, it would be an advantage to be able
to perform large time steps for the majority of the material (the coarse
parts) and smaller steps for the local patches where vertices could cause
instability. we call the concept of varying the timestep local adaptive
optimization of time step (LAOTS), in which ``local'' either refers to spatial
locality (as described above) or time locality where the time step is global
for all vertices but varies during the simulation. We will in this report
study possible approaches for application of LAOTS. In order to keep the focus
on possible LAOTS methods and remove unneccessary complexities we will study
the simple 1 dimensional simulation of a mass spring system and try to apply
LAOTS for it. Afterwards we will shortly discuss the application of our gained
knowledge to the domain of hyper elastic material simulation and simulations
in general.

@@

With no prerequisites within the field of mass simulations we will first
consider the 1D problem of simulating a mass spring system using the explicit
(forward) Euler method.

Second we will propose and test out three different methods for applying LAOTS
to the simulation.

Third and last we will discussion the results and how we
could apply our knowledge of LAOTS for hyper elastic material.

\subsection{1D mass spring system}
Mass spring systems are systems simulating a set of vertices interconnected by
springs in various ways to obtain desirable physical properties. The motion of
the springs and vertices are then simulated as a function of the time elapsed.
Examples of application of mass spring systems are for simulation of cloth or
hair.

Our mass spring system consists of a set $V$ of vertices and a set $S$ of
springs. Each vertex $v \in V$ has a specified position $p_v$, mass $m_v$, force
$F_v$
and velocity $u_v$ and each spring $s$ has two endpoints vertices $p_{sf} \in V$
and $p_{st} \in V \setminus p_{sf}$, a damping factor $d_s$, a stiffness
constant $K_s$ and a rest length $l_{s0}$. Combined all these variables and
constants describe the state of the mass spring system at each timestep during a
simulation.

As each vertex $v$ is connected to 1 or 2 springs, its motion can be calculated by
the total force $F_v$ applied by the springs it is connected to. The spring
force is calculated using Hooke's Law and adding damping [TODO: Add reference (wiki):
    \url{http://en.wikipedia.org/wiki/Harmonic_oscillator}]:
\begin{align}
    \label{eq:hooke}
    F_{h,d} &= -k_d x_d = -k_d (l_d - l_{0,d}) \\
    \label{eq:damping}
    F_{C,d} &= -C_d \frac{dx_d}{dt} = - C_d \frac{\left(v_{left} - v_{right}\right)
\Delta l}{\abs{\Delta l}}
\end{align}
in which $F_h$ is the spring force of Hooke's Law, $F_c$ is the damping force,
$k$ is the spring constant, $x$ is the difference between the current length of
the spring and the spring rest length, $c$ is the damping constant and
$\frac{dx}{dt}$ is the velocity of the vertex $v$.

@@

\subsection{Spatial LAOTS}
Atomic update
